---
title: "Final Project: Exploration of Social Media Use"
author: "Riya Belani and Moyalo Ifebajo"
date: "2025-06-26"
params:
  number: 6
  purpose: "Final project for PUBH 6199: Visualizing Data with R"
format: 
  html:
    toc: true
    toc-location: right
    code-fold: true
    theme: flatly
    self-contained: true
execute:
  echo: true
  warning: false
  message: false
output-dir: docs
---


```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "figs/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3
)
```

## Introduction

_Briefly introduce your topic and research question._

Research Questions
1. 
2. 
3. 

## Data

_Describe your data source(s), including where it came from, the key variables, and any pre-processing steps._

```{r}
#| eval: false
# Load libraries and data
library(tidyverse)

data <- read_csv("app-data/train.csv")
glimpse(data)
```   

## Shiny app: Exploring Social Media Use

_If you choose to create a Shiny app, describe the app's purpose and functionality. Save your code inside the shiny-app/ folder, replace the URL below with your app's URL_

[Click here to open the interactive Shiny app](https://8rjtwv-riya-belani.shinyapps.io/shiny-app/)

## Interpretation

_Summarize your key findings. What do your visualizations reveal?_

## Limitations

_Discuss any data quality or design limitations._

## Conclusion

_Wrap up your project with a takeaway message or next steps._

## References

_Include any references to data sources, packages, or literature you used. If your visualization design is inspired by someone else's work, cite those too._

1. ChatGPT 4.0 Version for Troubleshooting
2. Shiny Assistant for UI ideas
3. Kaggle Data Set
4. Packages: 


